---
title: "Plot_ly Project Page - NOAA Analysis"
output: html_document
date: "11-1-2023"
---

## Problem 2 Work: Creating Dashboard Graphs using `plotly`

For Homework 4 I am using the NOAA Meteorology Data Repository that is labeled `ny_noaa` in the p8105.datasets package. 

This page will provide a step-by-step guide of the code and thought processing I used as I designed the visualizations that are on the [dashboard](dashboard.html) page.

```{r Libraries and Settings, message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(plotly)

theme_set(theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Receiving initial guidance from the Data Import Method that has been Standardized in class lessons that used NOAA data across state sites, before doing specific manipulations to `nyweather_df`. This dataset has snow and snowfall depth variables which were usually excluded in class.

For these exercises, I am focusing on the weather data for years **2006 - 2010** and will be filtering the dataset ahead of time.

```{r Data Import and Initial Cleaning}
data("ny_noaa")

nyweather_df = ny_noaa |> 
  janitor::clean_names() |> 
  filter(date >= "2006-01-01") |> 
  filter(date <= "2010-12-31") |> 
  mutate(tmax = as.numeric(tmax), tmin = as.numeric(tmin)) |> 
  mutate(
    tmin = (tmin/10),
    tmax = (tmax/10))

#Counting distinct weather station id's that are being used.
test_count = nyweather_df |> 
  group_by(id) |> 
  summarize(n_obs = n()) |> 
  arrange(desc(n_obs))

```

There are currently `r nrow(test_count)` distinct observational ID weather stations in this dataframe. Specifically **145 distinct stations have full 1826 data collection points** across this full 5 year time period of recorded data (beginning of 2006 to end of 2010, 365 x 5 = 1826).

A separate dataframe containing only full reported observational data for each day can be created through filtering here, and will be called `full_nyweather_df` Note however this does not guarantee missing values from stations that do not record a specific variable.

```{r}
full_nyweather_df = nyweather_df |> 
  group_by(id) |> 
  mutate(n_obs = n()) |> 
  filter(n_obs >= 1826) |> 
  ungroup() |> 
  select(everything(), -n_obs)
```

`full_nyweather_df` has `r nrow(full_nyweather_df)` observations across 145 station ids, containing `rncol(full_nyweather_df)` column variables before full manipulation.

To make data more manageable, a small dataset `weather_sites_df` was created containing 3 Sites of interest across the Geography of NY State, 2006 - 2010, which have a monitoring point every day from 2006-2010.

```{r}
weather_sites_df = full_nyweather_df |> 
   filter(id == "USW00014768" | id == "USW00014735" | id == "USW00094728") |> 
   mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NYC",
      USW00014735 = "Albany_Mid_NY",
      USW00014768 = "Rochester_Upstate_NY"),
    month = lubridate::floor_date(date, unit = "month")) |> 
  mutate(name = as.factor(name)) |> 
  mutate(name = ordered(name, c("CentralPark_NYC", "Albany_Mid_NY", "Rochester_Upstate_NY"))) |> 
  select(name, id, date, month, everything())
 
weather_sites_df
```

Analyses for Visualization are then conducted, mirroring exercises performed in class with the NOAA API data across 3 sites.

### Graph 1: Temperature Change Seasonality Across 3 Sites 

```{r}
temps = weather_sites_df |> 
  drop_na(tmax) |> 
  group_by(name, month)

plot_ly(temps, x = ~month, y = ~tmax, color = ~name, type = "box", colors = "viridis") |> 
  layout(title = 'Max Temperature across 3 NY Monitors, 2006 - 2010', xaxis = list(title = 'Month'),
         yaxis = list(title = 'Recorded Tmax Degrees C'), legend = list(title=list(text=' Monitoring Station')))

```

### Graph 2: Measuring Average Monthly Precipitation Overtime across 3 NY Sites, 2006 - 2010

```{r}
precip_plot = weather_sites_df |> 
  drop_na(prcp) |> 
  group_by(name, month) |>
  summarize(avg_prcp = mean(prcp))

plot_ly(precip_plot, x= ~month, y = ~avg_prcp, color = ~name, type = "scatter", mode = "markers") |>
  add_lines(line = list(shape = 'spline', smoothing = 0.75)) |> 
  layout(title = 'Average Monthly Precipitation across 3 NY Monitors, 2006 - 2010', xaxis = list(title = 'Month'), 
         yaxis = list(title = 'Average Monthly Precipitation'), legend = list(title=list(text=' Monitoring Station')))
```

### Graph 3: Total Weekly Snowfall Across 3 NY Sites, 2008 - 2010

```{r}
snowfall = weather_sites_df |>
  filter(date >= "2008-01-01") |> 
  mutate(week = lubridate::floor_date(date, unit = "week")) |> 
  select(name, date, month, week, snow) |> 
  drop_na(snow) |> 
  mutate(rank = min_rank(desc(snow))) |>
  arrange(rank)

plot_ly(snowfall, x= ~week, y = ~snow, color = ~name, type= "bar") |> 
  layout(title = 'Weekly Snowfall across 3 NY Monitors, 2008 - 2010', xaxis = list(title = 'Week'), 
         yaxis = list(title = 'Total Weekly Snowfall (mm)'), legend = list(title=list(text=' Monitoring Station')))
```
