---
title: "Plot_ly Project Page - NOAA Analysis"
output: html_document
date: "11-1-2023"
---

## Problem 2 Work: Creating Dashboard Graphs using `plotly`

For Homework 4 I am using the NOAA Meteorology Data Repository that is labeled `ny_noaa` in the p8105.datasets package. 

This page will provide a step-by-step guide of the code and thought processing I used as I designed the visualizations that are on the [dashboard](dashboard.html) page.

```{r Libraries and Settings, message = FALSE}
library(tidyverse)
library(p8105.datasets)
library(plotly)
library(ggridges)
library(patchwork)

theme_set(theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Receiving initial guidance from the Data Import Method that has been Standardized in class lessons that used NOAA data across state sites, before doing specific manipulations to `nyweather_df`. This dataset has snow and snowfall depth variables which were usually excluded in class.

For these exercises, I am focusing on the weather data for years **2006 - 2010** and will be filtering the dataset ahead of time.

```{r Data Import and Initial Cleaning}
data("ny_noaa")

nyweather_df = ny_noaa |> 
  janitor::clean_names() |> 
  filter(date >= "2006-01-01") |> 
  filter(date <= "2010-12-31") 

#Counting distinct weather station id's that are being used.
test_count = nyweather_df |> 
  group_by(id) |> 
  summarize(n_obs = n()) |> 
  arrange(desc(n_obs))

```

There are currently `r nrow(test_count)` distinct observational ID weather stations in this dataframe. Specifically **145 distinct stations have full 1826 data collection points** across this full 5 year time period of recorded data (beginning of 2006 to end of 2010, 365 x 5 = 1826).

A separate dataframe containing only full reported observational data for each day can be created through filtering here, and will be called `full_nyweather_df` Note however this does not guarantee missing values from stations that do not record a specific variable.

```{r}
full_nyweather_df = nyweather_df |> 
  group_by(id) |> 
  mutate(n_obs = n()) |> 
  filter(n_obs >= 1826) |> 
  ungroup() |> 
  select(everything(), -n_obs)
```

full_nyweather_df has `r nrow(full_nyweather_df)` observations across 145 station ids, containing `rncol(full_nyweather_df)` column variables before full manipulation.

For particular specific data set groupings - I will need to convert the `date` format into separated years, months, and days to do average calculations over time, I will use this code below. 
  separate(col = date,
           into = c("year", "month", "day"),
           sep = "-", 
           convert = TRUE)

USW00014768 = "Rochester_Upstate_NY"
USW00014735 = "Albany_Mid_NY"
USW00094728 = "CentralPark_NYC


### Graph 1: Measuring Average Monthly Precipitation Overtime across NY Sites, 2006 - 2010

```{r}
rnoaa::ghcnd_stations(precip_plot)

precip_plot = full_nyweather_df |> 
  drop_na(prcp) |> 
  group_by(id) |> 
  plot_ly(x = ~date, y = ~prcp,
          type = "scatter", colors = "viridis")

precip_plot
  

```

